Likelihood free inference of the halo model parameters with group and clustering statistics of the simulations of galaxy mock catalogs

### 2016-01-07 halo/text/abc.tex
text so far

\documentclass[12pt, preprint]{aastex}
\usepackage[breaklinks,colorlinks, urlcolor=blue,citecolor=blue,linkcolor=blue]{hyperref}
\usepackage{graphicx}	% For figures
\usepackage{natbib}	% For citep and citep
\usepackage{amsmath}	% for \iint
\usepackage{bm}
\usepackage[breaklinks]{hyperref}	% for blackboard bold numbers
\usepackage{hyperref}
\hypersetup{colorlinks}
\usepackage{color}
\usepackage{morefloats}
\definecolor{darkred}{rgb}{0.5,0,0}
\definecolor{darkgreen}{rgb}{0,0.5,0}
\definecolor{darkblue}{rgb}{0,0,0.5}
\hypersetup{ colorlinks,
linkcolor=darkblue,
filecolor=darkgreen,
urlcolor=darkred,
citecolor=darkblue }

@@

\begin{abstract}
{\bf other abstract for reference}\\
The Halo Occupation Distribution (HOD) model is a powerful prescription of how galaxies populate dark matter halos that has had much success in reproducing large scale galaxy clustering and understanding galaxy evolution within the context of the halo model. Constraints on the parameters of the HOD model have so far been imposed using the standard practice of Bayesian inference with likelihood estimation that assumes a functional (often times Gaussian) form of the likelihood. Likelihood free inference methods such as the Approximate Bayesian Computation (ABC) makes Bayesian inference possible without ever having to calculate the likelihood.

In this work, we present ABC in the context of constraining the parameters of the HOD model. For a mock galaxy catalog with known HOD parameters, we use ABC on measurements of a set of observables generated through {\it Halotools}, a publicly available ..., to infer tight constraints that reproduce the known HOD paramters. Since ABC does not require computing the likelihood, we include observables beyond galaxy clustering such as the galaxy number density and the galaxy group richness, for which accurately constructing a likelihood is intractable. We demonstrate that adding these measurements to our set of observables leads to stronger constraints on the HOD parameters, 
 specifically the ones that govern the satellite galaxy population. Moreover, the constraints we infer that reproduce the known HOD parameters illustrate the success of ABC in the context of constraining HOD parameters.

The Halo Occupation Distribution model is a powerful prescription of how galaxies populate 
dark matter halos that has had much success in reproducing large scale galaxy clustering and understanding
galaxy evolution within the context of the halo model. Constraints on the parameters of the HOD model have 
so far been inferred using a likelihood function that assumes a functional--Gaussian--form of the likelihood
and an analytic model for the galaxy clustering statistics. Likelihood free inference methods such as the Approximate Bayesian Computation (ABC) makes Bayesian inference possible without ever having to calculate the likelihood.

In this work, we use ABC in order to infer the HOD parameters by assuming a prior over the parameters, and a distance function between 
statistical summaries of the data and those of the simulations.
Using a mock galaxy catalog created with known HOD parameters, we obtain precise constraints on the HOD parameters that 
reproduces the summary statistics of the mock galaxy catalog accurately. 
Since ABC does not require computing the likelihood, we include observable beyond galaxy clustering such as 
the galaxy number density and the galaxy group multiplicity function. 
We demonstrate that adding these measurements leads to stronger constraints on the HOD parameters, specifically the parameters 
that control the population of the satellite galaxies.

@@

Furthermore, This sampling formalism permits us to sample the posterior of the model by simulating the data via a forward generative model that accounts for all the complications and uncertainties associated with the survey data that are difficult to take into account in approximating the likelihood function. In the context of the large galaxy surveys, these complications include masks, nontrivial survey geometry, and missing data as a result of fiber collision.

@@

In practice, this rejected sampling method is replaced by a more efficient procedure. Instead, we start by sampling from the prior with a relatively large threshold. In the consequent steps, we iteratively perform importance sampling from the previous samples, and lower the threshold adaptively.

This method requires a simulator or a forward model of the positions of galaxies given the model parameters. We use the publicly available code Halotools in order to populate the halo catalogs with galaxies brighter than a certain luminosity threshold.

\section{Method}

@@

\section{Forward modeling of the galaxy population with HALOTOOLS}

@@

\section{Mock data}
Plots showing 2pcf, gmf, ... of the mock data.
How we compute the covariance matrix from mocks...
Talk about error bars. Talk about poisson errors in group populations ...

\section{Results}
Show the posteriors...
Show evolution of the posterior errorbars with iteration. Show epsilon. Convice that we are converging ...
Draw samples from the posterior, populate the mock catalogs with those samples, and compute the summary statistcs with them 
and demonstrate that they match the data...

@@

\section{Discussion}
\todo{talk about cosmological infeence and why this is good first step for likelihood free inference in large scale ...}

@@

talk about writing down a likelihood for gmf, tpcf, ... , and the gaussianity assumption ....

>>>>

\documentclass[12pt, preprint]{aastex}
\usepackage[breaklinks,colorlinks, urlcolor=blue,citecolor=blue,linkcolor=blue]{hyperref}
\usepackage{graphicx}	% For figures
\usepackage{natbib}	% For citep and citep
\usepackage{amsmath}	% for \iint
\usepackage{bm}
\usepackage[breaklinks]{hyperref}	% for blackboard bold numbers
\usepackage{hyperref}
\hypersetup{colorlinks}
\usepackage{color}
\usepackage{morefloats}
\definecolor{darkred}{rgb}{0.5,0,0}
\definecolor{darkgreen}{rgb}{0,0.5,0}
\definecolor{darkblue}{rgb}{0,0,0.5}
\hypersetup{ colorlinks,
linkcolor=darkblue,
filecolor=darkgreen,
urlcolor=darkred,
citecolor=darkblue}

@@

\begin{abstract}
The Halo Occupation Distribution (HOD) model is a powerful prescription of how galaxies
populate dark matter halos that has had much success in reproducing large scale galaxy
clustering and understanding galaxy evolution within the context of the halo model.
Constraints on the parameters that describe the HOD model have so far been inferred using 
the standard practice of Bayesian inference with likelihood estimation that assumes a
functional Gaussian form of the likelihood.\todo{(Do we want to include analytic models for clustering statistics here?)} Likelihood free inference methods such as the
Approximate Bayesian Computation (ABC) makes Bayesian inference possible without ever 
having to calculate the likelihood.

In this work, we present ABC in the context of constraining the parameters of the HOD model.
For a mock galaxy catalog created with HOD parameters selected from literature
\todo{(Should we specify Zheng07?)}, we use ABC on measurements of a set of observables
generated through {\it Halotools}, a publicly available \todo{insert description about
forward modeling of galaxy positions}, to infer precise constraints that reproduce the 
known HOD parameters.

In this work, we use ABC in order to infer the HOD parameters by assuming a prior over the parameters, and a distance function between 
statistical summaries of the data and those of the simulations.
Using a mock galaxy catalog created with known HOD parameters, we obtain precise constraints on the HOD parameters that 
reproduces the summary statistics of the mock galaxy catalog accurately.

Furthermore, since ABC does not require computing the likelihood, we include observables
beyond galaxy clustering such as the galaxy number density and the galaxy group 
multiplicity function, for which accurately constructing a likelihood is intractable. We demonstrate that adding these measurements to our set of observables leads to stronger constraints on the HOD parameters, 
 specifically the ones that govern the satellite galaxy population. Moreover, the constraints we infer that reproduce the known HOD parameters illustrate the success of ABC in the context of constraining HOD parameters.

Since ABC does not require computing the likelihood, we include observable beyond galaxy clustering such as 
the galaxy number density and the galaxy group multiplicity function. 
We demonstrate that adding these measurements leads to stronger constraints on the HOD parameters, specifically the parameters 
that control the population of the satellite galaxies.

@@

In practice, the rejected sampling method is replaced by a more efficient adaptive importance sampling operation called Population Monte Carlo (PMC). Initially, PMC rejects the draws from prior pdf with a relatively large threshold. In the consequent steps, importance sampling from the previous samples that passed the distance criteria is performed, and the threshold is updated adaptively. That is, after each iteration, the threshold is lowered according to the distribution of the distances.

Furthermore, This sampling formalism permits us to sample the posterior of the model by simulating the data via a forward generative model that accounts for all the complications and uncertainties associated with the survey data that are difficult to take into account in approximating the likelihood function. In the context of the large galaxy surveys, these complications may include masks, nontrivial survey geometry, and missing data as a result of fiber collision.

@@

The calculations regarding populating the dark matter halos, computing the clustering statistics and finding groups in the galaxy mock catalogs are performed using ({\fontfamily{cmss}\selectfont Halotools}) which is an open-source package tailored for modeling galaxy-halo connection in large scale structure modeling ({\fontfamily{cmss}\selectfont http://halotools.readthedocs.org}).

This paper is structured as follows. In Section \ref{sec:method},
we discuss the components of the Halo occupation modeling, and the ABC sampling method. We also give a brief overview of the summary statistics used in this investigation. In Section \ref{sec:halo} we give a brief overview of the forward modeling of the galaxy mock catalogs using Halotools as well as generating a mock galaxy catalog as data. In Section \ref{sec:results} we explore the results of our parameter inference using different summary statistics. Finally, we discuss and conclude in Section \ref{sec:discussion}.

\section{Method}\label{sec:method}

@@

\section{Forward modeling of the galaxy population with HALOTOOLS}\label{sec:halo}

@@

\subsection{Mock data}
Plots showing 2pcf, gmf, ... of the mock data.
How we compute the covariance matrix from mocks...
Talk about error bars. Talk about poisson errors in group populations ...

\section{Results}\label{sec:results}
Show the posteriors...
Show evolution of the posterior errorbars with iteration. Show epsilon. Convice that we are converging ...
Draw samples from the posterior, populate the mock catalogs with those samples, and compute the summary statistcs with them 
and demonstrate that they match the data...

@@

\section{Discussion}\label{sec:discussion}
\todo{talk about cosmological inference and why this is good first step for likelihood free inference in large scale ...}

@@

talk about writing down a likelihood for gmf, tpcf, ... , and the Gaussianity assumption ....
